Albacore North ASPIC assessment
===========================================================================

```{r}
# sets Rmd options globally
opts_chunk$set(comment   =NA, 
               fig.width =8,           
               fig.height=5.5, 
               warning   =FALSE, 
               message   =FALSE,
               eval      =FALSE,
               echo      =TRUE)
```

Initialisation Code
---------------------------------------------------------------------
```{r}
library(aspic)
dirMy="/home/laurie/Desktop/flr/git/aspic/inst/etc"
asp=aspic(paste(dirMy,"albn.inp",sep="/"))
asp=fit(asp)

key=data.frame(
  name   =c("Troll Composite CPUE","JLL Old","JLL Modern","CT Old","CT Modern"), 
  series=c("I","I","II","I","II"),
  flag  =c("OT","JA","JA","CT","CT"),
  gear  =c("TR","LL","LL","LL","LL"))

dimnames(key)[[1]]=c("Troll Composite CPUE","JLL Old","JLL Modern","CT Old","CT Modern") 

wts=t(array(c(1,1,1,1,1,
              1,0,0,0,0,
              0,1,1,0,0,
              0,0,0,1,1),c(5,4),list(name=key$name,Scenario=1:4)))

save(asp,key,wts,file=paste(dirMy,"Data/albn.RData",sep="/"))
```

CPUE
----------------------------------------------------------------
```{r}
cpue=subset(diags(asp),!is.na(obs))[,c("year","name","obs")]
ggplot(aes(year,obs,group=name,col=name),data=cpue)+
  geom_point()+
  stat_smooth()+
  theme_ms(legend.position="bottom")
```

```{r}
library(gam)
gm  =gam(log(obs)~lo(year)+name,data=cpue)
cpue=data.frame(cpue,gam=predict(gm),gamRsdl=residuals(gm))
scl =coefficients(gm)[3:9]
names(scl)=substr(names(scl),5,nchar(names(scl)))
cpue=transform(cpue,scl=scl[as.character(name)])
cpue[is.na(cpue$scl),"scl"]=0

cpue=cbind(cpue,key[cpue$name,])[,-2]
cpue$name=factor(cpue$name, levels=c("Troll Composite CPUE","JLL Old","JLL Modern","CT Old","CT Modern"))
ggplot(cpue)+ geom_line(aes(year,exp(gam)),col="red")  +
              geom_smooth(aes(year,obs),se=FALSE)      +           
              geom_point( aes(year,obs,col=name))       +
              facet_wrap(~name,ncol=1,scale="free_y")  +
              theme_ms(legend.position="none")         +
              xlab("Year") + ylab("Index")
```


```{r}
uMat=ddply(cpue,.(name),transform, obs=stdz(obs))
uMat=cast(uMat,year~name,value="obs")
uMat=uMat[apply(uMat,1,function(x) !all(is.na(x))),]

pM=plotmatrix(uMat[,-1])
pM$layers[[2]]=NULL
mns=ddply(subset(pM$data,!(is.na(x) & !is.na(y))),.(xvar,yvar), function(x) mean(x$y,na.rm=T))
pM+geom_hline(aes(yintercept=V1),data=mns,col="red") +
   geom_smooth(method="lm",se=F)  + 
   theme(legend.position="bottom")                   +
   xlab("Index")+ylab("Index")      
```


```{r, fig.height=8}
cr=cor(uMat[,-1],use="pairwise.complete.obs")
dimnames(cr)=list(gsub("_"," ",names(uMat)[-1]),gsub("_"," ",names(uMat)[-1]))
cr[is.na(cr)]=0
corrplot(cr,diag=F,order="hclust",addrect=2)  +          
             theme(legend.position="bottom")  
```


Fit
----------------------------------------------------------------
```{r}
plot(asp)
```


Residual Analysis
----------------------------------------------------------------
```{r}
rsdl=diags(asp)
```

**Observed CPUE verses fitted, blue line is a linear resgression fitted to points, black the y=x line.}**
```{r}
ggplot(ddply(rsdl, .(name), with, data.frame(obs=stdz(obs),hat=stdz(hat))))+
          geom_abline(aes(0,1))                         +
          geom_point( aes(obs,hat))                     +
          stat_smooth(aes(obs,hat),method="lm",fill="blue", alpha=0.1)       +
          facet_wrap(~name,ncol=3,scale="free")                              +
          theme_ms(legend.position="bottom")            +
          xlab("Fitted") + ylab("Observed")
```

**Residuals by year, with lowess smoother and SEs.**
```{r}
dat=ddply(subset(rsdl,!is.na(residual)),.(name), transform, residual=stdz(residual,na.rm=T))
ggplot(aes(year,residual),data=dat)  +
  geom_hline(aes(yintercept=0))      +
  geom_point()                       +
  stat_smooth(,method="loess",se=T,fill="blue", alpha=0.1)  +
  facet_wrap(~name,scale="free_x",ncol=2)  
```

**Plot of autocorrelation, i.e. $residual_{t+1}$ verses $residual_{t}$.**
```{r}
ggplot(rsdl)                                              +
  geom_point( aes(residual,residualLag))                  +
  stat_smooth(aes(residual,residualLag),method="lm",se=T,fill="blue", alpha=0.1)      +
  geom_hline(aes(yintercept=0))                           +
  facet_wrap(~name,scale="free",ncol=3)                   +
  xlab(expression(Residual[t])) + 
  ylab(expression(Residual[t+1])) +
  theme_ms(legend.position="bottom")  
```

**Quantile-quantile plot to compare residual distribution with the normal distribution.**
```{r}
ggplot(rsdl)                                           +
  geom_point( aes(qqx,qqy))                            +
  stat_smooth(aes(qqx,qqHat),method="lm",se=T,fill="blue", alpha=0.1)         +
  facet_wrap(~name)                                    +
  theme_ms(legend.position="bottom")                  
```

**Plot of residuals against fitted value, to check variance relationship.**
```{r}
ggplot(aes(hat, residual),data=subset(rsdl,!is.na(hat) & !is.na(residual)))   +
  geom_hline(aes(yintercept=0))         +
  geom_point()                          +
  stat_smooth(method="loess",span=.9,fill="blue", alpha=0.1)   +
  facet_wrap(~name,scale="free",ncol=3) 
```

Profiles to check
----------------------------------------------------------------
```{r,cache=FALSE}
asp@control[,"val"]=asp@params[,1]

prfl=rbind(cbind(Param="K",  profile(asp,which="k",  range=seq(0.80,1.20,length.out=21))))
```

```{r}
ggplot(prfl)+geom_line(aes(k,rss))
```


Sensitivity Analysis
----------------------------------------------------------------
```{r}
albB0=profile(asp,which="b0", range=seq(0.80,1.20,length.out=21),fn=function(x) x)

plot(albB0,fn=list("Stock"=function(x) stock(x)/bmsy(x), "Harvest"=function(x) harvest(x)/fmsy(x)))
```

Bootstrap
----------------------------------------------------------------
```{r}
albBoot=boot(asp)

plot(albBoot,fn=list("Stock"  =function(x) stock(x)/bmsy(x), 
                     "Harvest"=function(x) harvest(x)/fmsy(x)))+
  geom_hline(aes(yintercept=1),col="red")
```

```{r}
library(kobe)
kobePhase(model.frame(mcf(FLQuants("stock"  =stock(  albBoot)/bmsy(albBoot),
                                   "harvest"=harvest(albBoot)/fmsy(albBoot)))))+
  geom_point(aes(stock,harvest))
```
